---
title: "EDA"
output: github_document
date: "2023-11-27"
---

```{r}
library (tinytex)
```


# import data
```{r}
library (arrow)
library (tidyverse)
parquet_file <- "C:/Users/fsdel/Documents/2. Learning & Knowledge/1. Data Analytics/Data Science - Masters/Courses/Fall 23/IDS/Final Project/Cleaned Files/Energy Data divided by day by category.parquet"
df_house <- read_parquet(parquet_file)


file_path <- "C:/Users/fsdel/Documents/2. Learning & Knowledge/1. Data Analytics/Data Science - Masters/Courses/Fall 23/IDS/Final Project/Cleaned Files/Weather Data Cleaned.csv"
df_weather <- read_csv(file_path)

file_path <- "C:/Users/fsdel/Documents/2. Learning & Knowledge/1. Data Analytics/Data Science - Masters/Courses/Fall 23/IDS/Final Project/Cleaned Files/Static House Data Cleaned.csv"
df_static <- read_csv (file_path)


parquet_file <- "C:/Users/fsdel/Documents/2. Learning & Knowledge/1. Data Analytics/Data Science - Masters/Courses/Fall 23/IDS/Final Project/Cleaned Files/Energy Data divided by month by category.parquet"
df_house_by_month<-read_parquet(parquet_file)
```

# Data preparation: Merge all data by day

## Aggregate weather data by county and by day

```{r}
df_weather_mean_by_day<- df_weather %>%
  group_by(county, date_time = lubridate::floor_date(date_time, "day")) %>%
  summarise (mean_temp = mean (Dry.Bulb.Temperature...C.))
```




# which month is the coldest and hottest?

```{r}
# ----
# Which month is the hottest
  # Create a line chart to view which is the hottest month but is not clear if june
  # or july
ggplot(df_weather)+ 
  aes(x=date_time, y=Dry.Bulb.Temperature...C.) +
  xlab("Time of year")+
  ylab ("Weather Temperature")+
  geom_point()+
  ggtitle("Temperature by month")
  # doing an agreggation to determine which is the hottest month. We can see that
  # july is the month with the highest min temperature and max temperature compared 
  # to june. The coldest month is January.
df_weather %>%
  group_by(month) %>% 
  summarise(mean = mean (Dry.Bulb.Temperature...C.),
            min = min (Dry.Bulb.Temperature...C.),
            max = max (Dry.Bulb.Temperature...C.),
            median = median (Dry.Bulb.Temperature...C.))
```



# Does any variable is dependent on the temperature?
Explore correlations with temperature:
Some variables are correlated with the temperature like AC or heating but others are not

```{r}
parquet_file <- "C:/Users/fsdel/Documents/2. Learning & Knowledge/1. Data Analytics/Data Science - Masters/Courses/Fall 23/IDS/Final Project/EDA/All data merged by day.parquet"

df_merge<-read_parquet(parquet_file)

# code to create a graph and see the correlation between two variables
df_merge %>% ggplot () + aes (x= in.sqft, y=total_consumption) +geom_point() + geom_smooth(method="lm", se=TRUE)

# Identify numeric columns
num_cols <- sapply(df_merge, is.numeric)

sum (num_cols)

# DO the correlation matrix for temperature

cor_matrix_temp <- cor (df_merge[, num_cols],df_merge$mean_temp)
cor_matrix_temp
```

# Does any variable is dependent on the house size?
Explore correlation with SQFT
Some variables are correlated with the SQFT like plug or light but others are not

```{r}
# DO the correlation matrix for sqft
cor_matrix_sqt <- cor (df_merge[, num_cols],df_merge$in.sqft)

cor_matrix_sqt
```


